---
title: "Some Simple SPOOKY Data Analysis"
author: "Alek Anichowski"
date: "February 2, 2018"
output:
  html_document: default
  pdf_document: default
---

## Setup the libraries

```{r}
packages.used <- c("ggplot2", "dplyr", "tibble", "tidyr",  "stringr", "tidytext", "topicmodels", "wordcloud", "ggridges","forcats")

# check packages that need to be installed.
packages.needed <- setdiff(packages.used, intersect(installed.packages()[,1], packages.used))

# install additional packages
if(length(packages.needed) > 0) {
  install.packages(packages.needed, dependencies = TRUE, repos = 'http://cran.us.r-project.org')
}

library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(stringr)
library(tidytext)
library(topicmodels)
library(wordcloud)
library(ggridges)
library(forcats)

source("../lib/multiplot.R")
```

## Read in the data

```{r}
spooky <- read.csv('../data/spooky.csv', as.is = TRUE)
```

## Data Cleaning

```{r}
# Make a table with one word per row and remove `stop words` (i.e. the common words).
#1-gram
spooky_wrd <- unnest_tokens(spooky, word, text)
spooky_wrd <- anti_join(spooky_wrd, stop_words, by = "word")

#bigrams
spooky_bigrams <- unnest_tokens(spooky, bigram, text, token = "ngrams", n = 2)
bigrams_separated <- separate(spooky_bigrams,bigram, c("word1", "word2"), sep = " ")
bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
bigrams_united <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ")

#trigrams
spooky_trigrams <- unnest_tokens(spooky, trigram, text, token = "ngrams", n = 3)
trigrams_separated <- separate(spooky_trigrams,trigram, c("word1", "word2", "word3"), sep = " ")
trigrams_filtered <- trigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  filter(!word3 %in% stop_words$word)
trigrams_united <- trigrams_filtered %>%
  unite(trigram, word1, word2, word3, sep = " ")
```

## Unigram word counts

We see here that there are some common words to be found in the work of all the authors, "time", "life", "night", "eyes". MWS uses "love" a lot, and HPL uses "strange"

```{r}
# Unigram
author_words <- count(group_by(spooky_wrd, word, author))

all_words    <- rename(count(group_by(spooky_wrd, word)), all = n)

author_words <- left_join(author_words, all_words, by = "word")
author_words <- arrange(author_words, desc(all))
author_words <- ungroup(head(author_words, 81))
  
ggplot(author_words) +
  geom_col(aes(reorder(word, all, FUN = min), n, fill = author)) +
  xlab(NULL) +
  coord_flip() +
  facet_wrap(~ author) +
  theme(legend.position = "none")
```

## Bigram Analysis

Here we take a look at the common bigrams that occur in the documents. Unlike the unigrams, there seem to be a lot less overlap between the authors and the bigrams they use - "short time" being a noteworthy one, maybe used to create a sense of urgency in all the texts. 

Although there aren't a lot of common bigrams, we can use them to differentiate the authors. 
  - Names like "lord raymond" or "madame lalande" or "dr armitage" are usually specific to only 1 author. 
  - EAP and HPL are both fond of laughter, but they write it differently "ha ha" vs. "heh heh"
  - MWS deals with nature/animalistic themes - "fellow creatures", "native country" and "natural philosophy" are frequent 
  - On the other hand, EAP likes material things - "chess players", "main compartment", and "tea pot"
  - HPL describes locations a lot, like "shunned house", "ancient house", or "tempest mountain"

```{r}
# Bigram
author_bigrams <- count(group_by(bigrams_united, bigram, author))
all_bigrams    <- rename(count(group_by(bigrams_united, bigram)), all = n)
author_bigrams <- left_join(author_bigrams, all_bigrams, by = "bigram")
author_bigrams <- arrange(author_bigrams, desc(all))
author_bigrams <- author_bigrams[author_bigrams$all >10,]

bigrams_frequency <- count(bigrams_united, bigram, author)
bigrams_tf_idf    <- bind_tf_idf(bigrams_frequency, bigram, author, n)
bigrams_tf_idf    <- arrange(bigrams_tf_idf, desc(tf_idf))
bigrams_tf_idf    <- mutate(bigrams_tf_idf, bigram = factor(bigram, levels = rev(unique(bigram))))
bigrams_tf_idf_30 <- top_n(bigrams_tf_idf, 30, tf_idf)

  
ggplot(author_bigrams) +
  geom_col(aes(reorder(bigram, all, FUN = min), n, fill = author)) +
  xlab(NULL) +
  coord_flip() +
  facet_wrap(~ author) +
  theme(legend.position = "none") +
  labs(title = "Bigram Counts by Author")

ggplot(bigrams_tf_idf_30) +
  geom_col(aes(bigram, tf_idf, fill = author)) +
  labs(x = NULL, y = "TF-IDF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Bigram TF-IDF")
```
## Trigram Analysis

It seems that only EAP and HPL have trigrams that they like to repeat, at least in this dataset.
  - The most common trigrams are names, specific to each author, like "charles le sorcier" or "moreland clapham lee"
  - EAP still likes chess players, in fact they are usually "automaton chess players"
  - We see laughter again, still differentiated by spelling "ha ha ha" vs "heh heh heh". HPL's "heh" sounds more snide
  
```{r}
# Trigram
author_trigrams <- count(group_by(trigrams_united, trigram, author))
all_trigrams    <- rename(count(group_by(trigrams_united, trigram)), all = n)
author_trigrams <- left_join(author_trigrams, all_trigrams, by = "trigram")
author_trigrams <- ungroup(author_trigrams[author_trigrams$all>3,])

trigrams_frequency <- count(trigrams_united, trigram, author)
trigrams_tf_idf    <- bind_tf_idf(trigrams_frequency, trigram, author, n)
trigrams_tf_idf    <- arrange(trigrams_tf_idf, desc(tf_idf))
trigrams_tf_idf    <- mutate(trigrams_tf_idf, trigram = factor(trigram, levels = rev(unique(trigram))))
trigrams_tf_idf_30 <- top_n(trigrams_tf_idf, 20, tf_idf)
  
ggplot(author_trigrams) +
  geom_col(aes(reorder(trigram, all, FUN = min), n, fill = author)) +
  xlab(NULL) +
  coord_flip() +
  facet_wrap(~ author) +
  theme(legend.position = "none") +
  labs(title = "Trigram Frequencies")

ggplot(trigrams_tf_idf_30) +
  geom_col(aes(trigram, tf_idf, fill = author)) +
  labs(x = NULL, y = "TF-IDF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Trigram TF-IDF")
```

## Genders

The gender of pronouns also differentiates the authors. Looking at the counts of gendered pronouns, MWS uses far more female words than the other two-in fact, she uses far more pronouns in general, male and female. Looking at the frequency, this analysis holds up-MWS is balanced in her use of female to male characters (around 0.4 female vs 0.6 male). EAP is more male skewed, while HPL has a large bias towards using male characters.

```{r}

gender_wrd = unnest_tokens(spooky, word, text) %>%
        mutate(male = (word == 'he'| word == 'him'| word == 'his'|word =='man' |word == 'gentleman'|word == 'boy')) %>%
        mutate(female = (word == 'she' | word == 'her'| word == 'hers'|word =='woman'|word == 'lady'|word == 'girl')) %>%
        unite(sex, male, female) %>%
        mutate(sex = fct_recode(as.factor(sex), male = "TRUE_FALSE", female = "FALSE_TRUE", na = "FALSE_FALSE")) %>% 
        filter(sex != "na")

gender_frequency <- count(gender_wrd, sex, author)
gender_tf_idf    <- bind_tf_idf(gender_frequency, sex, author, n)

ggplot(gender_wrd,aes(sex, fill = author)) +
  labs(x = "Genders") +
  geom_bar(position = "dodge") +
  labs(title = "Gender Counts")

ggplot(gender_tf_idf) +
  geom_col(aes(sex, tf, fill = author),position = "dodge") +
  labs(x = NULL, y = "TF-IDF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Gender Frequency")

```

## Singular vs Plural Pronouns

Looking at singular vs plural pronouns, we see that all authors use the singular more than the plural, at almost the same rate in fact. MWS does tend to use the singular a bit more (80/20 split instead of 70/30 like EAP and HPL)

```{r}
pronoun_wrd = unnest_tokens(spooky, word, text) %>%
        mutate(single = (word == 'he'| word == 'him'| word == 'his'|word =='she' |word == 'her'|word == 'hers' |word =='I'|word =='me'|word =='my'|word=='mine')) %>%
        mutate(plural = (word == 'we'|word=='us'|word=='our'|word=='ours'|word=='they'|word=='them'|word=='their'|word=='theirs')) %>%
        unite(pronoun, single, plural) %>%
        mutate(pronoun = fct_recode(as.factor(pronoun), single = "TRUE_FALSE", plural = "FALSE_TRUE", na = "FALSE_FALSE")) %>% 
        filter(pronoun != "na")

pronoun_frequency <- count(pronoun_wrd, pronoun, author)
pronoun_tf_idf    <- bind_tf_idf(pronoun_frequency, pronoun, author, n)

ggplot(pronoun_wrd,aes(pronoun, fill = author)) +
  labs(x = "Pronouns") +
  geom_bar(position = "dodge") +
  labs(title = "Pronoun Counts(Single/Plural)")

ggplot(pronoun_tf_idf) +
  geom_col(aes(pronoun, tf, fill = author),position = "dodge") +
  labs(x = NULL, y = "TF-IDF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Pronoun Frequency(Single/Plural)")
```