---
title: "Some Simple SPOOKY Data Analysis"
author: "Alek Anichowski"
date: "February 2, 2018"
output:
  html_document: default
  pdf_document: default
---

## Setup the libraries

```{r}
packages.used <- c("ggplot2", "dplyr", "tibble", "tidyr",  "stringr", "tidytext", "topicmodels", "wordcloud", "ggridges","forcats")

# check packages that need to be installed.
packages.needed <- setdiff(packages.used, intersect(installed.packages()[,1], packages.used))

# install additional packages
if(length(packages.needed) > 0) {
  install.packages(packages.needed, dependencies = TRUE, repos = 'http://cran.us.r-project.org')
}

library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(stringr)
library(tidytext)
library(topicmodels)
library(wordcloud)
library(ggridges)
library(forcats)

source("../lib/multiplot.R")
```

## Read in the data

```{r}
spooky <- read.csv('../data/spooky.csv', as.is = TRUE)
```

## Data Cleaning

```{r}
# Make a table with one word per row and remove `stop words` (i.e. the common words).
#1-gram
spooky_wrd <- unnest_tokens(spooky, word, text)
spooky_wrd <- anti_join(spooky_wrd, stop_words, by = "word")

#bigrams
spooky_bigrams <- unnest_tokens(spooky, bigram, text, token = "ngrams", n = 2)
bigrams_separated <- separate(spooky_bigrams,bigram, c("word1", "word2"), sep = " ")
bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
bigrams_united <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ")

#trigrams
spooky_trigrams <- unnest_tokens(spooky, trigram, text, token = "ngrams", n = 3)
trigrams_separated <- separate(spooky_trigrams,trigram, c("word1", "word2", "word3"), sep = " ")
trigrams_filtered <- trigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  filter(!word3 %in% stop_words$word)
trigrams_united <- trigrams_filtered %>%
  unite(trigram, word1, word2, word3, sep = " ")
```

## Unigram word counts

We see here that there are some common words to be found in the work of all the authors, "time", "life", "night", "eyes". MWS uses "love" a lot, and HPL uses "strange"

```{r}
# Unigram
author_words <- count(group_by(spooky_wrd, word, author))

all_words    <- rename(count(group_by(spooky_wrd, word)), all = n)

author_words <- left_join(author_words, all_words, by = "word")
author_words <- arrange(author_words, desc(all))
author_words <- ungroup(head(author_words, 81))

ggplot(author_words) +
  geom_col(aes(reorder(word, all, FUN = min), n, fill = author)) +
  xlab(NULL) +
  coord_flip() +
  facet_wrap(~ author) +
  theme(legend.position = "none") +
  labs(title = 'Unigram Counts')
ggsave('unigram_counts.png',path='../figs')
```

## Bigram Analysis

Here we take a look at the common bigrams that occur in the documents. Unlike the unigrams, there seem to be a lot less overlap between the authors and the bigrams they use - "short time" being a noteworthy one, maybe used to create a sense of urgency in all the texts. 

Although there aren't a lot of common bigrams, we can use them to differentiate the authors. 
  - Names like "lord raymond" or "madame lalande" or "dr armitage" are usually specific to only 1 author. 
  - EAP and HPL are both fond of laughter, but they write it differently "ha ha" vs. "heh heh"
  - MWS deals with nature/animalistic themes - "fellow creatures", "native country" and "natural philosophy" are frequent 
  - On the other hand, EAP likes material things - "chess players", "main compartment", and "tea pot"
  - HPL describes locations a lot, like "shunned house", "ancient house", or "tempest mountain"

```{r}
# Bigram
author_bigrams <- count(group_by(bigrams_united, bigram, author))
all_bigrams    <- rename(count(group_by(bigrams_united, bigram)), all = n)
author_bigrams <- left_join(author_bigrams, all_bigrams, by = "bigram")
author_bigrams <- arrange(author_bigrams, desc(all))
author_bigrams <- author_bigrams[author_bigrams$all >10,]

bigrams_frequency <- count(bigrams_united, bigram, author)
bigrams_tf_idf    <- bind_tf_idf(bigrams_frequency, bigram, author, n)
bigrams_tf_idf    <- arrange(bigrams_tf_idf, desc(tf_idf))
bigrams_tf_idf    <- mutate(bigrams_tf_idf, bigram = factor(bigram, levels = rev(unique(bigram))))
bigrams_tf_idf_30 <- top_n(bigrams_tf_idf, 30, tf_idf)

  
ggplot(author_bigrams) +
  geom_col(aes(reorder(bigram, all, FUN = min), n, fill = author)) +
  xlab(NULL) +
  coord_flip() +
  facet_wrap(~ author) +
  theme(legend.position = "none") +
  labs(title = "Bigram Counts by Author")
ggsave('bigram_counts.png',path='../figs')

ggplot(bigrams_tf_idf_30) +
  geom_col(aes(bigram, tf_idf, fill = author)) +
  labs(x = NULL, y = "TF-IDF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Bigram TF-IDF")
ggsave('bigram_tfidf.png',path='../figs')
```
## Trigram Analysis

It seems that only EAP and HPL have trigrams that they like to repeat, at least in this dataset.
  - The most common trigrams are names, specific to each author, like "charles le sorcier" or "moreland clapham lee"
  - EAP still likes chess players, in fact they are usually "automaton chess players"
  - We see laughter again, still differentiated by spelling "ha ha ha" vs "heh heh heh". HPL's "heh" sounds more snide
  
```{r}
# Trigram
author_trigrams <- count(group_by(trigrams_united, trigram, author))
all_trigrams    <- rename(count(group_by(trigrams_united, trigram)), all = n)
author_trigrams <- left_join(author_trigrams, all_trigrams, by = "trigram")
author_trigrams <- ungroup(author_trigrams[author_trigrams$all>3,])

trigrams_frequency <- count(trigrams_united, trigram, author)
trigrams_tf_idf    <- bind_tf_idf(trigrams_frequency, trigram, author, n)
trigrams_tf_idf    <- arrange(trigrams_tf_idf, desc(tf_idf))
trigrams_tf_idf    <- mutate(trigrams_tf_idf, trigram = factor(trigram, levels = rev(unique(trigram))))
trigrams_tf_idf_30 <- top_n(trigrams_tf_idf, 20, tf_idf)
  
ggplot(author_trigrams) +
  geom_col(aes(reorder(trigram, all, FUN = min), n, fill = author)) +
  xlab(NULL) +
  coord_flip() +
  facet_wrap(~ author) +
  theme(legend.position = "none") +
  labs(title = "Trigram Frequencies")
ggsave('trigram_counts.png',path='../figs')

ggplot(trigrams_tf_idf_30) +
  geom_col(aes(trigram, tf_idf, fill = author)) +
  labs(x = NULL, y = "TF-IDF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Trigram TF-IDF")
ggsave('trigram_tfidf.png',path='../figs')
```

## Genders

The gender of pronouns also differentiates the authors. Looking at the counts of gendered pronouns, MWS uses far more female words than the other two-in fact, she uses far more pronouns in general, male and female. Looking at the frequency, this analysis holds up-MWS is balanced in her use of female to male characters (around 0.4 female vs 0.6 male). EAP is more male skewed, while HPL has a large bias towards using male characters.

```{r}

gender_wrd = unnest_tokens(spooky, word, text) %>%
        mutate(male = (word == 'he'| word == 'him'| word == 'his'|word =='man' |word == 'gentleman'|word == 'boy'|word=='himself')) %>%
        mutate(female = (word == 'she' | word == 'her'| word == 'hers'|word =='woman'|word == 'lady'|word == 'girl'|word=='herself')) %>%
        unite(sex, male, female) %>%
        mutate(sex = fct_recode(as.factor(sex), male = "TRUE_FALSE", female = "FALSE_TRUE", na = "FALSE_FALSE")) %>% 
        filter(sex != "na")

gender_frequency <- count(gender_wrd, sex, author)
gender_tf_idf    <- bind_tf_idf(gender_frequency, sex, author, n)

ggplot(gender_tf_idf) +
  geom_col(aes(sex, tf, fill = author),position = "dodge") +
  labs(x = NULL, y = "TF-IDF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Gender Frequency")

ggsave('gender_frequency.png',path='../figs')


```
## Gender Sentiment Analysis

Here we isolate the sentences that mention male or females, then do sentiment analysis on the words in that sentence. First we exclude sentences that mention both male and female genders, as we are interested in the difference betweent the two. 

Interestingly, since there are less female sentences overall(1837 female vs 5176 male), the 525 shared sentences(with both men and women in them) constitute a larger percentage of the female sentences. 28% of the female sentences also had men in them, while only 10% of the male sentences had women, meaning the females are mentioned along with men with higher frequency than men were mentioned with women. 

Going on to the sentiment analysis, the counts are not that useful, since there are so many more male sentences, so the counts are also biased. The graph shows the frequency of sentiments across the authors, specifically the difference between the male frequency and female frequency of a certain sentiment. A positive difference means that the male sentences were more associated with this sentiment, and negative female sentences. A few trends emerge:

More significant differences:
  -EAP trusts men more than women, this seems to be a major trait of his sentiments towards men.
  -EAP and MWS associate joy and sadness with women more than men, so they are more emotional in the corpus.
  -EAP's men are more positive, while the women more negative - the other author's don't have this difference between gendered pos and neg.
  
Less significant differences:
  -Only in HPL are men more sad than women-in EAP and MWS sadness is associated with women more.
  -HPL's women are also more suprised, angry and disgusted than the men.
  -MWS's treatment of the genders is fairly balanced
```{r}
male_wrd <- gender_wrd %>%
  group_by(id) %>%
  filter(sex =='male')
male_groups <- unique(male_wrd[,1])

female_wrd <- gender_wrd %>%
  group_by(id) %>%
  filter(sex =='female')
female_groups <- unique(female_wrd[,1])

male_only <- anti_join(male_groups,female_groups)
female_only <- anti_join(female_groups,male_groups)
male_lines <- spooky %>%
  filter(id %in% male_only$id)
female_lines <- spooky %>%
  filter(id %in% female_only$id)

get_sentiments('nrc')
male_wrd <- unnest_tokens(male_lines, word, text)
female_wrd <- unnest_tokens(female_lines, word, text)

male_sentiments <- inner_join(male_wrd, get_sentiments('nrc'), by = "word")
female_sentiments <- inner_join(female_wrd, get_sentiments('nrc'), by = "word")

male_sentiment_frequency <- count(male_sentiments, sentiment, author)
male_tf_idf    <- bind_tf_idf(male_sentiment_frequency, sentiment, author, n)

female_sentiment_frequency <- count(female_sentiments, sentiment, author)
female_tf_idf    <- bind_tf_idf(female_sentiment_frequency, sentiment, author, n)

male_tf_idf$diff <- male_tf_idf$tf-female_tf_idf$tf

ggplot(male_tf_idf) +
  geom_col(aes(sentiment, diff, fill = author),position = "dodge") +
  labs(x = NULL, y = "Difference in TF values") +
  theme(axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Male - Female Sentiment Frequency Difference")

ggsave('gender_sentiments.png',path='../figs')

```
## Singular vs Plural Pronouns

Looking at singular vs plural pronouns, we see that all authors use the singular more than the plural, at almost the same rate in fact. MWS does tend to use the singular a bit more (80/20 split instead of 70/30 like EAP and HPL)

```{r}
pronoun_wrd = unnest_tokens(spooky, word, text) %>%
        mutate(single = (word == 'he'| word == 'him'| word == 'his'|word =='she' |word == 'her'|word == 'hers' |word =='I'|word =='me'|word =='my'|word=='mine')) %>%
        mutate(plural = (word == 'we'|word=='us'|word=='our'|word=='ours'|word=='they'|word=='them'|word=='their'|word=='theirs')) %>%
        unite(pronoun, single, plural) %>%
        mutate(pronoun = fct_recode(as.factor(pronoun), single = "TRUE_FALSE", plural = "FALSE_TRUE", na = "FALSE_FALSE")) %>% 
        filter(pronoun != "na")

pronoun_frequency <- count(pronoun_wrd, pronoun, author)
pronoun_tf_idf    <- bind_tf_idf(pronoun_frequency, pronoun, author, n)

ggplot(pronoun_tf_idf) +
  geom_col(aes(pronoun, tf, fill = author),position = "dodge") +
  labs(x = NULL, y = "TF values") +
  theme(legend.position = "top", axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Pronoun Frequency(Single/Plural)")

ggsave('pronoun_frequency.png',path='../figs')

```
## Single/Plural Pronoun Sentiment Analysis
Like with gender, we take sentences that exclusively use single or plural pronouns and analyze the difference in frequency of various sentiments in those sentences.

  -In EAP the plural sentences are more positive than the single, but for HPL it's reversed, single is more positive than plural.
  -EAP trusts groups more, wheras HPL and MWS trust the individual.
  -HPL and MWS associate fear with the plural more than the single pronoun.
```{r}
single_wrd <- pronoun_wrd %>%
  group_by(id) %>%
  filter(pronoun =='single')
single_groups <- unique(single_wrd[,1])

plural_wrd <- pronoun_wrd %>%
  group_by(id) %>%
  filter(pronoun =='plural')
plural_groups <- unique(plural_wrd[,1])

single_only <- anti_join(single_groups,plural_groups)
plural_only <- anti_join(plural_groups,single_groups)
single_lines <- spooky %>%
  filter(id %in% single_only$id)
plural_lines <- spooky %>%
  filter(id %in% plural_only$id)

get_sentiments('nrc')
single_wrd <- unnest_tokens(single_lines, word, text)
plural_wrd <- unnest_tokens(plural_lines, word, text)

single_sentiments <- inner_join(single_wrd, get_sentiments('nrc'), by = "word")
plural_sentiments <- inner_join(plural_wrd, get_sentiments('nrc'), by = "word")

single_sentiment_frequency <- count(single_sentiments, sentiment, author)
single_tf_idf    <- bind_tf_idf(single_sentiment_frequency, sentiment, author, n)

plural_sentiment_frequency <- count(plural_sentiments, sentiment, author)
plural_tf_idf    <- bind_tf_idf(plural_sentiment_frequency, sentiment, author, n)

single_tf_idf$diff <- single_tf_idf$tf-plural_tf_idf$tf

ggplot(single_tf_idf) +
  geom_col(aes(sentiment, diff, fill = author),position = "dodge") +
  labs(x = NULL, y = "Difference in TF values") +
  theme(axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Single - Plural Sentiment Frequency Difference")

ggsave('pronoun_sentiments.png',path='../figs')

```

## Light vs Dark

MWS and EAP use more light than dark imagery, and with MWS the difference is the greatest. HPL on the other hand uses more dark imagery in his works.

```{r}
lights_wrd = unnest_tokens(spooky, word, text) %>%
        mutate(light = (word == 'light'|word =='bright'|word=='luminous'|word=='rich'|word=='shiny'|word=='sunny'|word=='sun'|word=='burnished'|word=='flashing'|word=='glowing'|word=='shining'|word=='radiant'|word=='cloudless'|word=='polished'|word=='lustrous'|word=='white')) %>%
        mutate(dark = (word == 'dark'|word=='black'|word=='grey'|word=='gray'|word=='dim'|word=='darkened'|word=='dingy'|word=='dull'|word=='foggy'|word=='gloomy'|word=='shady'|word=='shadow'|word=='somber')) %>%
        unite(light, light, dark) %>%
        mutate(light = fct_recode(as.factor(light), light = "TRUE_FALSE", dark = "FALSE_TRUE", na = "FALSE_FALSE")) %>% 
        filter(light != "na")

light_frequency <- count(lights_wrd, light, author)
light_tf_idf    <- bind_tf_idf(light_frequency, light, author, n)

ggplot(light_tf_idf) +
  geom_col(aes(light, tf, fill = author),position = "dodge") +
  labs(x = NULL, y = "TF values") +
  theme(axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Light vs Dark Frequency")

ggsave('lights_frequency.png',path='../figs')

```

## Light/Dark Sentiments

When comparing light/dark sentiments, we pretty much get exactly what we hope for in terms of sentiment association, but there doesn't seem to be much differentiation between the authors.
  -anticipation, joy and positivity are associated with light.
  -negative and sadness are associated with dark.
  -EAP and MWS pair suprise with light, while HPL is unbiased
  -EAP has more disgust in his dark imagery
  -
```{r}
light_wrd <- lights_wrd %>%
  group_by(id) %>%
  filter(light =='light')
light_groups <- unique(light_wrd[,1])

dark_wrd <- lights_wrd %>%
  group_by(id) %>%
  filter(light =='dark')
dark_groups <- unique(dark_wrd[,1])

light_only <- anti_join(light_groups,dark_groups)
dark_only <- anti_join(dark_groups,light_groups)
light_lines <- spooky %>%
  filter(id %in% light_only$id)
dark_lines <- spooky %>%
  filter(id %in% dark_only$id)

get_sentiments('nrc')
light_wrd <- unnest_tokens(light_lines, word, text)
dark_wrd <- unnest_tokens(dark_lines, word, text)

light_sentiments <- inner_join(light_wrd, get_sentiments('nrc'), by = "word")
dark_sentiments <- inner_join(dark_wrd, get_sentiments('nrc'), by = "word")

light_sentiment_frequency <- count(light_sentiments, sentiment, author)
light_tf_idf    <- bind_tf_idf(light_sentiment_frequency, sentiment, author, n)

dark_sentiment_frequency <- count(dark_sentiments, sentiment, author)
dark_tf_idf    <- bind_tf_idf(dark_sentiment_frequency, sentiment, author, n)

light_tf_idf$diff <- light_tf_idf$tf-dark_tf_idf$tf

ggplot(light_tf_idf) +
  geom_col(aes(sentiment, diff, fill = author),position = "dodge") +
  labs(x = NULL, y = "Difference in TF values") +
  theme(axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = "Light - Dark Sentiment Frequency Difference")

ggsave('lights_sentiment.png',path='../figs')

```


